Fast solvers for partial differential equations \\ subject to inequalities

### 2023-11-04 talk-dms/slides.tex
a bit of progress

%\usepackage[svgnames]{xcolor}
\usepackage{hyperref}
\hypersetup{
    pdftitle = {Toward nonlinear multigrid for nonlinear and nonlocal variational inequalities},
    pdfauthor = {Ed Bueler and Patrick Farrell},
    pdfsubject = {},
    pdfkeywords = {}
}

@@

%\begin{frame}
%  \maketitle
%\end{frame}

@@

\begin{frame}{Outline}
  %\tableofcontents[hideallsubsections]
  \tableofcontents
\end{frame}

@@

@@

\bigskip
\only<1>{
\begin{itemize}
\item \emph{problem.} on a domain $\Omega \subset \RR^2$, find the displacement $u(x)$ of a membrane, with fixed value $u = g$ on $\partial \Omega$, above an \emph{obstacle} $\psi(x)$, which minimizes the elastic energy
    $$J(v) = \int_\Omega \frac{1}{2} |\grad v|^2 - f\, v$$
\item shown above: \quad $\Omega=(-2,2)^2$, $\psi(x)$ a hemisphere, $f(x)=0$
\end{itemize}

\phantom{x}
}
\only<2>{
\begin{itemize}
\item i.e.~constrained optimization over a convex \emph{admissible set}
	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
\item $J'(u)$ points directly into $\mathcal{K}$, the {\color{FireBrick} \emph{variational inequality} (VI)}:
    $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0 \quad \text{for all } v \in \mathcal{K}$$
\end{itemize}
}
\only<3>{
\begin{itemize}
\item the solution defines \emph{active} $A_u = \{u = \psi\}$ and \emph{inactive} $R_u = \{u> \psi\}$ subsets of $\Omega$, and a \emph{free boundary} $\Gamma_u=\partial R_u \cap \Omega$
\item naive strong form would pose the problem in terms of its solution:
\begin{align*}
-\grad^2 u &= f \quad \text{ on $R_u$} \\
u &= \psi \quad \text{ on $A_u$}
\end{align*}

@@

\phantom{x}
\end{itemize}
}
\only<4>{
\begin{itemize}
\item the \emph{complementarity problem} (CP) is meaningful as a strong form:
\begin{align*}
u - \psi &\ge 0 \\
-\grad^2 u - f &\ge 0 \\
(u - \psi)(-\grad^2 u - f) &= 0
\end{align*}
   \begin{itemize}
   \item[$\circ$] for optimization problems: \quad CP $=$ KKT conditions
   \end{itemize}

>>>>

%\usepackage[svgnames]{xcolor}
\usepackage{hyperref}
\hypersetup{
    pdftitle = {Fast solvers for PDEs subject to inequalities},
    pdfauthor = {Ed Bueler and Patrick Farrell},
    pdfsubject = {},
    pdfkeywords = {}
}

@@

@@

\begin{frame}{Outline}
  \tableofcontents[hideallsubsections]
  %\tableofcontents
\end{frame}

@@

\begin{center}
\includegraphics[width=0.65\textwidth]{../talk-oxford/images/obstacle65.pdf}
\end{center}

\only<1>{
\begin{itemize}
\item \emph{problem.} on a domain $\Omega \subset \RR^2$, find the displacement $u(x)$ of a membrane, with fixed value $u = g$ on $\partial \Omega$, above an \emph{obstacle} $\psi(x)$, which minimizes the elastic energy
    $$J(v) = \int_\Omega \frac{1}{2} |\grad v|^2 - f\, v$$
\item shown above: \quad $\Omega$ a square, $\psi(x)$ a hemisphere
\end{itemize}

\phantom{x}
}
\only<2>{
\begin{itemize}
\item i.e.~constrained optimization over an \emph{admissible set}
	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
\item $\mathcal{K}$ is a closed and convex subset of a Sobolev space:
	$$H^1(\Omega) = \left\{v \,:\, \int_\Omega |v|^2 + |\grad v|^2 < \infty\right\}$$
\end{itemize}
}
\end{frame}

@@

\bigskip
\only<1>{
\begin{itemize}
\item the solution defines \emph{active} $A_u = \{u = \psi\}$ and \emph{inactive} $R_u = \{u> \psi\}$ subsets of $\Omega$, and a \emph{free boundary} $\Gamma_u=\partial R_u \cap \Omega$
\item naive strong form would pose the problem in terms of its solution:
\begin{align*}
-\grad^2 u &= f \quad \text{ on $R_u$} \\
u &= \psi \quad \text{ on $A_u$}
\end{align*}

   \begin{itemize}
   \item[$\circ$] this is not acceptable \dots
   \end{itemize}
\end{itemize}
}
\only<2>{
\begin{itemize}
\item the \emph{complementarity problem} (CP) is meaningful as a strong form:
\begin{align*}
u - \psi &\ge 0 \\
-\grad^2 u - f &\ge 0 \\
(u - \psi)(-\grad^2 u - f) &= 0
\end{align*}

   \begin{itemize}
   \item[$\circ$] CP $=$ (KKT conditions in $\infty$-dimensions)
   \end{itemize}

@@

\phantom{x}
\end{itemize}
}
\only<3>{
\begin{itemize}
\item FIXME i.e.~constrained optimization over a convex \emph{admissible set}
	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
\item $J'(u)$ points directly into $\mathcal{K}$, the {\color{FireBrick} \emph{variational inequality} (VI)}:
    $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0 \quad \text{for all } v \in \mathcal{K}$$
\end{itemize}
}
\end{frame}
